---
title: "`r params$title`"
author: "`r params$author`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document:
    toc: true
    toc_depth: 2
    highlight: tango
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
params:
  title: "Kernel Analysis Report"
  author: ""
  data_file: ""
  include_doctor: true
  include_summary: true
  include_profile: true
  include_plots: true
  data: NULL
  doctor_findings: NULL
  fix_log: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE,
  fig.width = 8, 
  fig.height = 5,
  dpi = 150
)
library(ggplot2)
library(dplyr)
library(knitr)

df <- params$data
findings <- params$doctor_findings
fixes <- params$fix_log

kernel_blue <- "#2C5A8F"
kernel_dark <- "#1B3D66"

theme_kernel_report <- function(base_size = 12) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title = element_text(color = kernel_dark, face = "bold", size = base_size + 2),
      axis.title = element_text(color = kernel_dark),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#E8EDF2")
    )
}
```

---

# Executive Summary

This report was generated by **Kernel** — a free, open-source statistical analysis platform.

| Metric | Value |
|--------|-------|
| Observations | `r format(nrow(df), big.mark = ",")` |
| Variables | `r ncol(df)` |
| Numeric | `r sum(sapply(df, is.numeric))` |
| Categorical | `r sum(sapply(df, function(x) is.factor(x) || is.character(x)))` |
| Missing values | `r format(sum(is.na(df)), big.mark = ",")` |
| Complete rows | `r format(sum(complete.cases(df)), big.mark = ",")` |
| Completeness | `r round((1 - sum(is.na(df)) / (nrow(df) * ncol(df))) * 100, 1)`% |

---

```{r doctor, eval=params$include_doctor && !is.null(findings) && is.data.frame(findings) && nrow(findings) > 0, results='asis'}
cat("# Data Quality Assessment\n\n")

n_crit <- sum(findings$severity == "critical")
n_warn <- sum(findings$severity == "warning")
n_info <- sum(findings$severity == "info")

cat(sprintf("The Variable Doctor identified **%d** issues:\n\n", nrow(findings)))
cat(sprintf("- **Critical:** %d\n", n_crit))
cat(sprintf("- **Warnings:** %d\n", n_warn))
cat(sprintf("- **Suggestions:** %d\n\n", n_info))

kable(findings[, c("variable", "severity", "issue", "detail")],
      col.names = c("Variable", "Severity", "Issue", "Details"),
      caption = "Variable Doctor Findings")
```

```{r fixes, eval=!is.null(fixes) && length(fixes) > 0, results='asis'}
cat("\n## Fixes Applied\n\n")
for (i in seq_along(fixes)) {
  cat(sprintf("%d. %s\n", i, fixes[i]))
}
```

---

```{r profile, eval=params$include_profile, results='asis'}
cat("# Column Profile\n\n")

prof <- data.frame(
  Variable = names(df),
  Type = sapply(df, function(x) {
    if (is.numeric(x)) "numeric"
    else if (is.factor(x)) "factor"
    else if (is.character(x)) "character"
    else if (is.logical(x)) "logical"
    else if (inherits(x, "Date")) "date"
    else class(x)[1]
  }),
  Missing = sapply(df, function(x) sum(is.na(x))),
  `Missing %` = round(sapply(df, function(x) mean(is.na(x)) * 100), 1),
  Unique = sapply(df, function(x) length(unique(x[!is.na(x)]))),
  check.names = FALSE
)

kable(prof, caption = "Column-by-column profile")
```

---

```{r numeric_summary, eval=params$include_summary, results='asis'}
cat("# Numeric Summary Statistics\n\n")

nums <- df[, sapply(df, is.numeric), drop = FALSE]

if (ncol(nums) > 0) {
  summ <- do.call(rbind, lapply(names(nums), function(nm) {
    x <- nums[[nm]]
    data.frame(
      Variable = nm,
      n = sum(!is.na(x)),
      Mean = round(mean(x, na.rm = TRUE), 4),
      SD = round(sd(x, na.rm = TRUE), 4),
      Min = round(min(x, na.rm = TRUE), 4),
      Q1 = round(quantile(x, 0.25, na.rm = TRUE), 4),
      Median = round(median(x, na.rm = TRUE), 4),
      Q3 = round(quantile(x, 0.75, na.rm = TRUE), 4),
      Max = round(max(x, na.rm = TRUE), 4),
      check.names = FALSE
    )
  }))
  kable(summ, caption = "Descriptive statistics for numeric variables", row.names = FALSE)
} else {
  cat("No numeric variables in the dataset.\n")
}
```

---

```{r distributions, eval=params$include_plots && ncol(df[, sapply(df, is.numeric), drop = FALSE]) > 0, results='asis', fig.height=4}
cat("# Distributions\n\n")

nums <- df[, sapply(df, is.numeric), drop = FALSE]
plot_vars <- names(nums)[1:min(8, ncol(nums))]

for (v in plot_vars) {
  cat(sprintf("\n## %s\n\n", v))
  p <- ggplot(df, aes(x = .data[[v]])) +
    geom_histogram(aes(y = after_stat(density)), bins = 30, 
                   fill = kernel_blue, alpha = 0.6, color = "white") +
    geom_density(linewidth = 1, color = kernel_dark) +
    labs(title = paste("Distribution of", v), x = v, y = "Density") +
    theme_kernel_report()
  print(p)
  cat("\n\n")
}
```

---

```{r categorical_summary, eval=params$include_summary, results='asis'}
cats <- df[, sapply(df, function(x) is.factor(x) || is.character(x)), drop = FALSE]

if (ncol(cats) > 0) {
  cat("# Categorical Variables\n\n")
  
  for (v in names(cats)[1:min(5, ncol(cats))]) {
    cat(sprintf("\n## %s\n\n", v))
    tab <- sort(table(cats[[v]], useNA = "ifany"), decreasing = TRUE)
    out <- data.frame(
      Level = names(tab),
      Count = as.integer(tab),
      Proportion = round(as.numeric(tab) / sum(tab), 3)
    )
    print(kable(head(out, 10), caption = paste("Top levels of", v)))
    cat("\n")
  }
}
```

---

```{r correlation, eval=params$include_plots, results='asis', fig.height=6}
nums <- df[, sapply(df, is.numeric), drop = FALSE]

if (ncol(nums) >= 2) {
  cat("# Correlation Analysis\n\n")
  
  cor_vars <- names(nums)[1:min(10, ncol(nums))]
  cor_mat <- cor(nums[, cor_vars, drop = FALSE], use = "pairwise.complete.obs")
  
  # Build long format without reshape2
  cor_long <- data.frame(
    Var1 = rep(rownames(cor_mat), ncol(cor_mat)),
    Var2 = rep(colnames(cor_mat), each = nrow(cor_mat)),
    value = as.vector(cor_mat),
    stringsAsFactors = FALSE
  )
  cor_long$Var1 <- factor(cor_long$Var1, levels = rownames(cor_mat))
  cor_long$Var2 <- factor(cor_long$Var2, levels = rev(colnames(cor_mat)))
  
  p <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), size = 3) +
    scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", 
                         midpoint = 0, limits = c(-1, 1)) +
    labs(title = "Correlation Matrix", x = NULL, y = NULL) +
    theme_kernel_report() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank())
  print(p)
  
  cat("\n\n")
  kable(round(cor_mat, 3), caption = "Pairwise correlations")
}
```

---

# Appendix

## Session Information

- **Generated by:** Kernel ([kernelstats.com](https://kernelstats.com))
- **Date:** `r format(Sys.time(), '%B %d, %Y at %I:%M %p')`
- **R version:** `r R.version.string`
- **Data file:** `r params$data_file`

---

*This report was automatically generated by Kernel — The core of your analysis.* 🌱
