---
title: "Analytical Workbench Report"
subtitle: "Parameterized report template for reproducible EDA, categorical analysis, GLMs, and DOE"
author: "Analytical Workbench Pro"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    theme: flatly
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
params:
  dataset_path: "data.csv"
  dataset_label: "Uploaded dataset"
  assoc_x: null
  assoc_y: null
  model_formula: null
  doe_formula: null
---

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)

read_any <- function(path) {
  ext <- tolower(tools::file_ext(path))
  if (ext == "csv") return(readr::read_csv(path, show_col_types = FALSE))
  if (ext %in% c("xlsx", "xls")) return(readxl::read_excel(path))
  if (ext == "tsv") return(readr::read_tsv(path, show_col_types = FALSE))
  if (ext == "txt") return(readr::read_delim(path, delim = ",", show_col_types = FALSE))
  stop("Unsupported file type for template. Extend read_any() as needed.")
}

df <- read_any(params$dataset_path)
num_vars <- names(df)[vapply(df, is.numeric, logical(1))]
cat_vars <- names(df)[vapply(df, function(x) is.factor(x) || is.character(x) || is.logical(x), logical(1))]
```

# Executive summary

**Dataset:** `r params$dataset_label`  
**Rows:** `r nrow(df)`  
**Columns:** `r ncol(df)`  
**Numeric variables:** `r length(num_vars)`  
**Categorical-like variables:** `r length(cat_vars)`

# Data profile

```{r}
profile <- tibble(
  variable = names(df),
  type = vapply(df, function(x) paste(class(x), collapse = "/"), character(1)),
  missing_n = vapply(df, function(x) sum(is.na(x)), numeric(1)),
  missing_pct = round(vapply(df, function(x) mean(is.na(x)) * 100, numeric(1)), 2),
  distinct_n = vapply(df, function(x) dplyr::n_distinct(x, na.rm = TRUE), numeric(1))
)
profile
```

## Missingness

```{r}
ggplot(profile, aes(x = reorder(variable, missing_pct), y = missing_pct)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL, y = "Missing (%)", title = "Missingness by variable") +
  theme_minimal(base_size = 12)
```

# Univariate review

```{r}
if (length(num_vars)) {
  first_num <- num_vars[1]
  ggplot(df, aes(x = .data[[first_num]])) +
    geom_histogram(bins = 30, color = "white") +
    labs(title = paste("Histogram of", first_num), x = first_num, y = "Count") +
    theme_minimal(base_size = 12)
}
```

```{r}
if (length(cat_vars)) {
  first_cat <- cat_vars[1]
  ggplot(df, aes(x = .data[[first_cat]])) +
    geom_bar() +
    coord_flip() +
    labs(title = paste("Bar chart of", first_cat), x = first_cat, y = "Count") +
    theme_minimal(base_size = 12)
}
```

# Categorical association

```{r}
if (!is.null(params$assoc_x) && !is.null(params$assoc_y) && all(c(params$assoc_x, params$assoc_y) %in% names(df))) {
  tbl <- table(df[[params$assoc_x]], df[[params$assoc_y]])
  tbl
  chisq.test(tbl)
}
```

# Regression / GLM

```{r}
if (!is.null(params$model_formula)) {
  model_formula <- as.formula(params$model_formula)
  fit <- lm(model_formula, data = df)
  summary(fit)
}
```

# DOE analysis

```{r}
if (!is.null(params$doe_formula)) {
  doe_formula <- as.formula(params$doe_formula)
  doe_fit <- lm(doe_formula, data = df)
  anova(doe_fit)
}
```

# Notes for extension

- Replace the placeholder `lm()` call with `glm()`, `nnet::multinom()`, `ordinal::clm()`, `MASS::glm.nb()`, or `pscl::zeroinfl()` depending on the selected model family.
- Add `FactoMineR` sections for PCA/MCA/FAMD when those packages are available.
- Add DOE-generation code chunks when you want the report to include generated designs, alias structures, or response-surface plots.
